# Comparing two proportions {-}

```{r inference-summary-tableRPT, echo=FALSE, message=FALSE, warning=FALSE}
# Original at https://docs.google.com/spreadsheets/d/1QkOpnBGqOXGyJjwqx1T2O5G5D72wWGfWlPyufOgtkk4/edit#gid=0
library(dplyr)
library(readr)
read_csv("data/ch9_summary_table - Sheet1.csv", na = "") %>% 
  kable(
    caption = "Scenarios of sample statisitics for inference (Table 1 repeated)", 
    booktabs = TRUE
  )
```

In the previous sections we have considered Scenario 2 in Table 1 (reproduced here), i.e. constructing a confidence interval for a single population mean.  Often, however, interest lies in comparing two populations, e.g. by constructing a confidence interval for the difference in the population means, as in Scenario 4.  But it may be that the characteristic of interest is a population proportion (rather than a population mean) which is reflected in Scenarios 1 and 3.  In this section we will focus on Scenario 3, i.e. constructing a confidence interval for the difference in two population proportions.

Let's start with an example.  If you see someone else yawn, are you more likely to yawn? In an episode of the TV show [*Mythbusters*](https://en.wikipedia.org/wiki/MythBusters), they tested the myth that yawning is contagious. 

Fifty adults who thought they were being considered for an appearance on the show were interviewed by a show recruiter ("confederate") who either yawned or did not. Participants then sat by themselves in a large van and were asked to wait. While in the van, the Mythbusters watched via hidden camera to see if the unaware participants yawned. The data frame containing the results is available at `mythbusters_yawn` in the `moderndive` package. Let's check it out.

```{r}
mythbusters_yawn
```

- The participant ID is stored in the `subj` variable with values of 1 to 50.
- The `group` variable is either `"seed"` for when a confederate was trying to influence the participant or `"control"` if a confederate did not interact with the participant.
- The `yawn` variable is either `"yes"` if the participant yawned or `"no"` if the participant did not yawn.

We can use the `janitor` package to get a glimpse into this data in a table format:

```{r}
mythbusters_yawn %>% 
  tabyl(group, yawn) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns() # To show original counts
```

We are interested in comparing the proportion of those that yawned after seeing a `seed` versus those that yawned with no seed interaction (i.e. `control`). We'd like to see if the difference between these two proportions is significantly larger than 0. If so, we'd have evidence to support the claim that yawning is contagious based on this study.

In looking over this problem, we can take note of some important details to include in our `infer` pipeline:

- We are calling a "`success`" having a `yawn` value of `yes`.
- Our response variable will always correspond to the variable used in the `success` so the response variable is `yawn`.
- The explanatory variable is the other variable of interest here: `group`.


## Compute the point estimate {-}

We are examining the relationship between yawning and whether or not the participant saw someone yawn (`seed`) or not (`control`).

```{r error=TRUE}
mythbusters_yawn %>% 
  specify(formula = yawn ~ group)
```

Note that the `success` argument must be specified in situations such as this where the response variable has only two levels.

```{r}
mythbusters_yawn %>% 
  specify(formula = yawn ~ group, success = "yes")
```

We next want to calculate the statistic of interest for our sample. This corresponds to the difference in the proportion of successes.

```{r error=TRUE, warning=TRUE}
mythbusters_yawn %>% 
  specify(formula = yawn ~ group, success = "yes") %>% 
  calculate(stat = "diff in props")
```

We see another error here. To further check to make sure that R knows exactly what we are after, we need to provide the `order` in which R should subtract these proportions of successes. As the error message states, we'll want to put `"seed"` first after `c()` and then `"control"`: `order = c("seed", "control")`. Our point estimate is thus calculated:

```{r error=TRUE}
obs_diff <- mythbusters_yawn %>% 
  specify(formula = yawn ~ group, success = "yes") %>% 
  calculate(stat = "diff in props", order = c("seed", "control"))
obs_diff
```

This value represents the proportion of those that yawned after seeing a seed yawn (0.2941) minus the proportion of those that yawned with not seeing a seed (0.25).


## Bootstrap distribution {-}

Our next step in building a confidence interval is to create a bootstrap distribution of statistics (differences in proportions of successes). We saw how it works with a single variable in computing bootstrap means in the `pennies` example but we haven't yet worked with bootstrapping involving multiple variables, i.e. comparing two groups. 
In the `infer` package, bootstrapping with multiple variables means that each **row** is potentially resampled. Let's investigate this by looking at the first few rows of `mythbusters_yawn`:

```{r}
head(mythbusters_yawn)
```

When we bootstrap this data, we are potentially pulling the subject's readings multiple times. Thus, we could see the entries of `"seed"` for `group` and `"no"` for `yawn` together in a new row in a bootstrap sample. This is further seen by exploring the `sample_n()` function in `dplyr` on this smaller 6 row data frame comprised of `head(mythbusters_yawn)`. The `sample_n()` function can perform this bootstrapping procedure and is similar to the `rep_sample_n()` function in `infer`, except that it is not `rep`eated but rather only performs one sample with or without replacement.

```{r, echo=FALSE}
set.seed(2019)
```

```{r}
head(mythbusters_yawn) %>% 
  sample_n(size = 6, replace = TRUE)
```

We can see that in this bootstrap sample generated from the first six rows of `mythbusters_yawn`, we have some rows repeated. The same is true when we perform the `generate()` step in `infer` as done below.

```{r, echo=c(1)}
bootstrap_distribution <- mythbusters_yawn %>% 
  specify(formula = yawn ~ group, success = "yes") %>% 
  generate(reps = 1000) %>% 
  calculate(stat = "diff in props", order = c("seed", "control"))

bootstrap_distribution
```

```{r}
bootstrap_distribution %>% 
  visualize(bins = 20)
```

This distribution is roughly symmetric and bell-shaped but isn't quite there. Let's use the percentile-based method to compute a 95% confidence interval for the true difference in the proportion of those that yawn with and without a seed presented. The arguments are explicitly listed here but remember they are the defaults and simply `get_ci()` can be used.

```{r}
bootstrap_distribution %>% 
  get_ci(type = "percentile", level = 0.95)
```

```{r include=FALSE}
bootstrap_distribution %>% 
  get_ci(type = "percentile", level = 0.95) -> myth_ci
```

<br>

```{r MCQ4, echo=FALSE}
opts_Q4 <- sample(c("We have evidence that you are more likely to yawn after seeing someone yawn",
                    answer = "There is insufficient evidence that the probability of  yawning is related to  seeing someone yawn",
                    "We have evidence that you are less likely to yawn after seeing someone yawn"))
```

**Looking at the confidence interval shown here:**
`r longmcq(opts_Q4)`

<br>

Therefore, we are not sure which proportion is larger. Some of the bootstrap statistics showed the proportion without a seed to be higher and others showed the proportion with a seed to be higher. If the confidence interval was entirely above zero, we would be relatively sure (about "95% confident") that the seed group had a higher proportion of yawning than the control group.

Note that this all relates to the importance of denoting the `order` argument in the `calculate()` function. Since we specified `"seed"` and then `"control"` positive values for the statistic correspond to the `"seed"` proportion being higher, whereas negative values correspond to the `"control"` group being higher.

We, therefore, have evidence via this confidence interval suggesting that the conclusion from the Mythbusters show that "yawning is contagious" being "confirmed" is **not** statistically correct.

<br>
<br>


