--- 
title: "Data Analysis: Confidence Intervals and Model Parameter Selection"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
biblio-style: apalike
link-citations: yes
---

```{r setupCI, include=FALSE}
library(webexercises)
library(ggplot2)
library(dplyr)
library(moderndive)
library(ISLR)
library(skimr)
library(plotly)
library(tidyr)
library(datasets)
library(knitr)
library(janitor)
library(infer)
library(readr)
knitr::opts_chunk$set(comment = NA, warning = FALSE, message = FALSE)
```

```{r setupModelParam, include=FALSE}
library(ggplot2)
library(dplyr)
library(moderndive)
library(ISLR)
library(skimr)
library(plotly)
library(tidyr)
library(datasets)
library(knitr)
library(janitor)
library(infer)
library(readr)
library(broom)
library(gridExtra)
library(GGally) #Package to produce matrix of 'pairs' plots and more!

knitr::opts_chunk$set(comment = NA, warning = FALSE, message = FALSE)
```

# Introduction {-}

In previous weeks we have seen many examples of calculating *sample statistics* such as means, percentiles, standard deviations and regression coefficients.  These *sample statistics* are used as *point estimates* of *population parameters* which describe the *population* from which the *sample* of data was taken.  That last sentence assumes you're familiar with concepts and terminology about sampling (e.g. from the *Statistical Inference* course in 1st Semester) so here is a summary of some key terms:

1. **Population**: The population is a set of $N$ observations of interest.

2. **Population parameter**: A population parameter is a numerical summary value about the population. In most settings, this is a value that's unknown and you wish you knew it.

3. **Census**: An exhaustive enumeration/counting of all observations in the population in order to compute the population parameter's numerical value *exactly*.
    + When $N$ is small, a census is feasible. However, when $N$ is large, a census can get very expensive, either in terms of time, energy, or money. 

4. **Sampling**: Collecting a sample of size $n$ of observations from the population. Typically the sample size $n$ is much smaller than the population size $N$, thereby making sampling a much cheaper procedure than a census. 
    + It is important to remember that the lowercase $n$ corresponds to the sample size and uppercase $N$ corresponds to the population size, thus  $n \leq N$.

5. **Point estimates/sample statistics**: A summary statistic based on the sample of size $n$ that *estimates* the unknown population parameter.

6. **Representative sampling**: A sample is said be a *representative sample* if it "looks like the population". In other words, the sample's characteristics are a good representation of the population's characteristics.

7. **Generalizability**: We say a sample is *generalizable* if any results based on the sample can generalize to the population.

8. **Bias**: In a statistical sense, we say *bias* occurs if certain observations in a population have a higher chance of being sampled than others. We say a sampling procedure is *unbiased* if every observation in a population had an equal chance of being sampled. 

9. **Random sampling**: We say a sampling procedure is *random* if we sample randomly from the population in an unbiased fashion.

***

Now that you are familiar with RMarkdown, you are encouraged to collate your work in this tutorial in a RMarkdown file. Create a `.Rmd` file to load the following packages into R:

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(janitor)
library(moderndive)
library(infer)
```

**Note**: Additional information and examples can be found in [Chapter 9](https://moderndive.com/9-confidence-intervals.html) of [An Introduction to Statistical and Data Science via R](https://moderndive.com/index.html).

<br>
<br>






















